---
title: "main-functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{main-functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(measlesr)
```

### Load measles data        
**This function loads the measles yearly data provided by the WHO and tidytuesday.**      
Takes 0 parameters, it can be saved as an object, however all functions within the package automatically load the data.        

```{r}
#| label: load data 
load_data()
```

### Validate measles year           
**This function validates the users input to ensure only years are selected from the study.**      
This is a helper cleaning function used in the summary, plot and model functions. It validates user inputs and curates informative responses when user inputs are invalid.    
Takes one parameter:    
- year (numeric)    

```{r}
#| label: load data 
validate_measles_year(2015)
```


### Measles snapshot    
**This function returns the top or  bottom 10 countries by total measles cases in the defined study.**     
Takes two parameters:     
- Year (numeric - 2012-2025)    
- top (TRUE = top 10, FALSE = bottom 10)      

```{r}
#| label: country summary
measles_snapshot(2022, top = TRUE)
```

### Measles lab coverage        
**This function returns the top or  bottom 10 countries by proportion of lab confirmed to total measles cases.**    
Takes two parameters:     
- Year (numeric - 2012-2025)    
- top (TRUE = top 10, FALSE = bottom 10)      

```{r}
#| label: lab coverage 
lab_coverage(2022, top = TRUE)
```


```{r}
#| label: hex plots

#this function would be able to create a hex plot based on the given variables, and maybe a predetermined layout of title/captions
#two parameters, var1 and var2, both need to be numerical 
create_hex_plot(var1, var2)
```

```{r}
#| label: leaflet map
# this function would create a leaflet map based on any selected country-level 
# metric and any selected year. The function will color in every country on the
# world map (using a custom gradient color scheme specified by the user) based 
# on the selected metric.
plot_leaflet_map(data = cases_year,
                 year = 2023,
                 palette = "YlOrRd",
                 metric = cases_per_100k)

```
